---
title: "Final Project"
author: "S&DS 230/S&DS 530/ENV 757/PLSC 530"
subtitle: Due by 11:59am, Saturday, May 7, 2022
output:
  html_document: default
  word_document: default
  pdf_document: default
---

### Introduction: Background and Motivation

Understanding whether or not a country’s capital openness affects foreign direct investment can help policymakers better understand how to attract (or reduce) foreign investment.

### Goal of Analysis

Our questions of interest:

1. Among the variables selected for analysis, which ones have the greatest influence on gdp per capita?
2.

### Data

#### List of Variables You Actually Use

#### Data cleaning process

**1) Data Cleaning** - demonstrate use of find/replace, data cleaning, dealing with missing values, text character replacement, matching. It is ok if your data did not require much of this.

Our data includes yearly data across a variety of variables for 175 countries from 1970-2019. 

| Variable Names     | Description                                                                                                                        |
| ------------------ | ---------------------------------------------------------------------------------------------------------------------------------- |
| country_name       | Country name                                                                                                                       |
| COWcode            | Numerical code for country -- from Correlates of War Project                                                                       |
| iso3c              | Alphabetical country code from ISO standard                                                                                        |
| year               | Year of data                                                                                                                       |
| fdi_flows          | Yearly net inflows of foreign direct investment                                                                                    |
| fdi_per_gdp        | Yearly net inflows of foreign direct investment as percentage of GDP                                                               |
| gdp                | Gross domestic product                                                                                                             |
| gdppc              | Gross domestic product per capita                                                                                                  |
| trade_dependence   | Total trade as percentage of GDP                                                                                                   |
| inflation          | Inflation rate                                                                                                                     |
| nat_resoc_rents    | Percentage of GDP derived from natural resources                                                                                   |
| bits               | Number of bilateral investment traties in force in the country                                                                     |
| wto                | Shows whether the country is a WTO member                                                                                          |
| Polity             | Polity5 Democracy Index -- higher values == stronger democracy                                                                     |
| v2x_polyarchy      | Electoral democracy index -- higher values == stronger democratic institutions                                                     |
| v2x_rule           | Rule of law index -- higher values == stronger rule of law                                                                         |
| v2x_gender         | Women's political empowerment index -- higher values == greater political empowerment for women                                    |
| polcon             | Political Constraint Index (veto players) -- higher values indicate greater number of veto players                                 |
| ka_open            | Measure of capital account openness -- higher values == greater integration with global financial markets                          |
| alliance_US        | Shows whether the country is a member of the US alliance                                                                           |
| alliance_RU        | Shows whether the country is a member of the RU alliance                                                                           |
| conflict           | Shows whether the country is in an armed conflict with any other country                                                           |
| conflict_intensity | Shows the intensity of armed conflict in the country (minor or major)                                                              |
| conflict_type      | Shows the type of armed conflict in the country (no conflict, extrasystemic, interstate, intrastate, internationalized intrastate) |
| physint            | Measures extent to which a state violates physical integrity rights -- higher values == greater respect for these rights           |

```{r}
library(dplyr)
```

```{r}
fdi <- read.csv("fdi_data.csv")
```

#### Data cleaning process

To start, looking at the variables "wto", "alliance_US", "alliance_RU", "conflict", "conflict_intensity", and "conflict_type", these variables are categorical but are represented as numbers. We will convert these variables to categorical variables to make it easier to work with and understand the data.

```{r}
fdi$wto <- as.logical(fdi$wto)
fdi$alliance_US <- as.logical(fdi$alliance_US)
fdi$alliance_RU <- as.logical(fdi$alliance_RU)
fdi$conflict <- as.logical(fdi$conflict)
fdi$conflict_intensity <- ifelse(fdi$conflict_intensity == 1, "minor", "major")
fdi$conflict_type[fdi$conflict_type == 0] <- "no conflict"
fdi$conflict_type[fdi$conflict_type == 1] <- "extrasystemic"
fdi$conflict_type[fdi$conflict_type == 2] <- "interstate"
fdi$conflict_type[fdi$conflict_type == 3] <- "intrastate"
fdi$conflict_type[fdi$conflict_type == 4] <- "internationalized intrastate"
dim(fdi)
head(fdi)
```

**1)  Data Cleaning**

*Variable Names and their Descriptions*

Categorical variables:

* conflict | Shows whether the country is in an armed conflict with any other country 

  * Value of 1 indicates conflict
  
  * Value of 0 indicates no conflict

* conflict_intensity | Shows the intensity of armed conflict in the country

  * Value of 0 indicates no conflict
  
  * Value of 1 indicates minor conflict
  
  * Value of 2 indicates major conflict
  
* conflict_type | Shows the type of conflict

  * Value of 0 indicates no conflict
  
  * Value of 1 indicates extrasystemic conflict
  
  * Value of 2 indicates interstate conflict
  
  * Value of 3 indicates intrastate conflict
  
  * Value of 4 indicates internationalized intrastate
  
Continuous:

* gdppc: Gross domestic product per capita (in USD)

* trade_dependence: Total trade (imports + exports) as percentage of GDP

* v2x_polyarchy

  * Value from 0 - 1
  
  * Electoral democracy index where higher values == stronger democratic institutions

* v2x_rule

  * Value from 0 - 1
  
  * Rule of law index -- measures extent to which laws are fairly enforced -- higher values == stronger rule of law

* v2x_gender: 

  * Value from 0 - 1 
  
  * Women's political empowerment index -- higher values == greater political empowerment for women

* ka_open: 

  * Value from 0 - 1
  
  * Measure of capital account openness -- higher values == greater integration with global financial markets


#### Data cleaning process

To start, looking at the variables conflict", "conflict_intensity", and "conflict_type", these variables are categorical but are represented as numbers. We will convert these values to booleans/more descriptive values to make it easier to work with and understand the data.

```{r}
fdi$conflict <- as.logical(fdi$conflict)
fdi$conflict_intensity[fdi$conflict_intensity == 0] <- "None"
fdi$conflict_intensity[fdi$conflict_intensity == 1] <- "Minor"
fdi$conflict_intensity[fdi$conflict_intensity == 2] <- "Major"
fdi$conflict_type[fdi$conflict_type == 0] <- "No conflict"
fdi$conflict_type[fdi$conflict_type == 1] <- "Extrasystemic"
fdi$conflict_type[fdi$conflict_type == 2] <- "Interstate"
fdi$conflict_type[fdi$conflict_type == 3] <- "Intrastate"
fdi$conflict_type[fdi$conflict_type == 4] <- "Internationalized intrastate"  
```

To further narrow down our data for this project, we will only be looking at data from countries in 2019. 
```{r}
fdi_2019 <- fdi[fdi$year == 2019,]
dim(fdi_2019)
head(fdi_2019)
```

Variables we want to use:
gdppc
trade_dependence

wto

Use the following to predict gdppc?

- v2x_polyarchy
- v2x_rule
- v2x_gender
- ka_open

**2) Scraping More Info on S&P 500 Companies** - show appropriate use of at least ONE of each of the following – boxplot,
scatterplot (can be matrix plot), normal quantile plot (can be related to regression),
residual plots, histogram.

```{r}
# create scatterplot of gdp per capita vs. democracy index in 2019
plot(fdi_2019$gdppc, fdi_2019$v2x_polyarchy, xlab="GDP per capita", ylab="Democracy Index")
title("GDP per capita vs. Democracy Index in 2019")

# create scatterplot of gdp per capita vs. rule of law index in 2019
plot(fdi_2019$gdppc, fdi_2019$v2x_rule, xlab="GDP per capita", ylab="Rule of Law Index")
title("GDP per capita vs. Rule of Law Index in 2019")

# create scatterplot of gdp per capita vs. women's policital empowerment in 2019
plot(fdi_2019$gdppc, fdi_2019$v2x_gender, xlab="GDP per capita", ylab="Women's Political Empowerment Index")
title("GDP per capita vs. Women's Political Empowerment Index in 2019")

# create scatterplot of gdp per capita vs. measure of capital account openness

plot(fdi_2019$gdppc, fdi_2019$ka_open, xlab="GDP per capita", ylab="Capital Account Openness")
title("GDP per capita vs. Capital Account Openness in 2019")

```

This scatterplot does not convey a lot of useful information due to the outliers on the right side of the plot. To counteract this, we can replot this with the log of the GDP per capita.

```{r}
fdi_2019$log_gdppc <- log(fdi_2019$gdppc)
# create scatterplot of gdp per capita vs. democracy index in 2019
plot(fdi_2019$log_gdppc, fdi_2019$v2x_polyarchy, xlab="GDP per capita", ylab="Democracy Index")
title("Log of GDP per capita vs. Democracy Index in 2019")

# create scatterplot of gdp per capita vs. rule of law index in 2019
plot(fdi_2019$log_gdppc, fdi_2019$v2x_rule, xlab="GDP per capita", ylab="Rule of Law Index")
title("Log of GDP per capita vs. Rule of Law Index in 2019")

# create scatterplot of gdp per capita vs. women's policital empowerment in 2019
plot(fdi_2019$log_gdppc, fdi_2019$v2x_gender, xlab="GDP per capita", ylab="Women's Political Empowerment Index")
title("Log of GDP per capita vs. Women's Political Empowerment Index in 2019")

# create scatterplot of gdp per capita vs. measure of capital account openness
plot(fdi_2019$log_gdppc, fdi_2019$ka_open, xlab="GDP per capita", ylab="Capital Account Openness")
title("Log of GDP per capita vs. Capital Account Openness in 2019")
```

After logging the gdppc, it is a bit easier to see that there is a slight positive correlation with gdppc for the democracy index and the rule of law index. Plotting the GDP per capita with the women's empowerment index does not show a significant correlation in any way.

**3) Basic Tests** - t-test, correlation, AND ability to create bootstrap confidence interval for
either a t-test or a correlation.

### Correlation

We chose to look at the correlation between FDI Per GDP and GDP Per Capita and found that out of the continuous indicator variables these had a weak correlation of 0.046.

```{r, echo = FALSE}
plot(jitter(fdi_2019$gdppc), jitter(fdi_2019$fdi_per_gdp), col = "green", xlab = "gdppc", ylab = "fdi_per_gdp")
mtext(paste("Correlation =", round(cor(fdi_2019$gdppc, fdi_2019$fdi_per_gdp, use = "complete.obs"), 3)), cex = 1.2, line = 0)
mtext("gdppc vs. fdi_per_gdp", cex = 1.2, line = 1)
abline(lm(fdi_2019$fdi_per_gdp ~ fdi_2019$gdppc), col = "orange")
```

```{r}
#Calculate the correlation between Conflict Intensity and GDP per Capita
(cor5 <- round(cor(fdi_2019$gdppc, fdi_2019$conflict_intensity, use = "pairwise.complete.obs"),3))
```

### T-Test

Using a t-test, we analyzed if there was a difference in means between alliance_US and alliance_RU. We found that there was a significant difference in means between the two variables.

```{r}
# Create a dataframe with just two columns for target variables
gddpc_ttest_data <- data.frame(fdi_2019$conflict, fdi_2019$gdppc)

# Remove rows with missing values in any column of data frame
gddpc_ttest_data<- gddpc_ttest_data[complete.cases(gddpc_ttest_data),]

colnames(gddpc_ttest_data) <- c("conflict", "gdppc")
(ttest <- t.test(gddpc_ttest_data$gdppc ~ gddpc_ttest_data$conflict, conf.level = .99))
```

Because the p-value is 0.000676, at an alpha of 0.05 we would conclude that there is a statistically significant difference in GDP per capita between countries that are currently in armed conflict and those that are not.

We then performed a bootstrapped 95% confidence interval:

```{r, echo=FALSE}
set.seed(230)

N <- 10000
bootstrapped_conflict <- rep(NA, N)
for (i in 1:N) {
  bootstrapped_conflict[i] <- mean(sample(gddpc_ttest_data$gdppc[gddpc_ttest_data$conflict == TRUE],
                   sum(gddpc_ttest_data$conflict == TRUE), replace = TRUE))
}
qLeft <- quantile(bootstrapped_conflict, c(0.025, 0.975))
paste0("The 95% bootstrapped confidence interval for GDP per capita of countries with armed conflict is (", qLeft[1], ", ", qLeft[2], ")")

bottstrapped_no_conflict <- rep(NA, N)
for (i in 1:N) {
  bottstrapped_no_conflict[i] <- mean(sample(gddpc_ttest_data$gdppc[gddpc_ttest_data$conflict == FALSE],
                   sum(gddpc_ttest_data$conflict == FALSE), replace = TRUE))
}
qRight <- quantile(bottstrapped_no_conflict, c(0.025, 0.975))
paste0("The 95% bootstrapped confidence interval for GDP per capita of countries with no armed conflict is (", qRight[1], ", ", qRight[2], ")")
```

The bootstrapped CIs indicate the same conclusion the t-test made which is that there is a significant difference in GDP per Capita between countries that are in armed conflicts versus those that are not.

**4) Permutation Test** - include at least one.

```{r, echo=FALSE}
set.seed(230)

N <- 10000
diffvals <- rep(NA, N)
for (i in 1:N) {
  fakeconflict <- sample(gddpc_ttest_data$conflict)
  diffvals[i] <- mean(gddpc_ttest_data$gdppc[fakeconflict == TRUE]) -
    mean(gddpc_ttest_data$gdppc[fakeconflict == FALSE])
}

actualdiff <- mean(gddpc_ttest_data$gdppc[gddpc_ttest_data$conflict == TRUE])  -
  mean(gddpc_ttest_data$gdppc[gddpc_ttest_data$conflict == FALSE])

hist(diffvals,
     col = "yellow",
     main = "Permuted Sample Mean Diff in gdppc by conflict status",
     xlab = "Difference in gdppc",
     breaks = 50)
abline(v = actualdiff, col = "blue", lwd = 3)
text(actualdiff-.01, 300, paste("Actual Diff in Means =", round(actualdiff, 4)), srt = 90)

paste("P-value:", mean(abs(diffvals) >= abs(actualdiff)))
```

The permuted sample mean difference in gdppc by party group suggests that the difference in the sample mean is approximately 0. The p-value of 0.7405 suggests that we fail to reject the null hypothesis and that at a significance level of 0.05 we conclude that the mean improvement in runtimes is not statistically significantly different between political groups.

**5) Multiple Regression** - use either backwards stepwise regression or some form of best
subsets regression. Should include residual plots. A GLM with a mix of continuous and
categorical predictors is fine here.

**6) AT LEAST ONE OF THE FOLLOWING TECHNIQUES** - ANOVA, ANCOVA, Logistic
Regression, Multinomial Regression, OR data scraping off a website.

```{r}
fdi_anova <- fdi_2019[c("gdppc", "conflict_intensity", "conflict_type")]
fdi_anova <- fdi_anova[complete.cases(fdi_anova),]

```

#### Conclusion and Summary
