---
title: "Final Project"
author: "S&DS 230/S&DS 530/ENV 757/PLSC 530"
subtitle: Due by 11:59am, Saturday, May 7, 2022
output:
  html_document: default
  word_document: default
  pdf_document: default
---

### Introduction: Background and Motivation

Understanding whether or not a country’s capital openness affects foreign direct investment can help policymakers better understand how to attract (or reduce) foreign investment.

### Goal of Analysis

Our questions of interest:

1. Among the variables selected for analysis, which ones have the greatest influence on gdp per capita?
2.

### Data

#### List of Variables You Actually Use

#### Data cleaning process

**1) Data Cleaning** - demonstrate use of find/replace, data cleaning, dealing with missing
values, text character replacement, matching. It is ok if your data did not require much of
this.

Variables we want to use:
gdppc
trade_dependence

wto

Use the following to predict gdppc?

- polity
- v2x_polyarchy
- v2x_rule
- v2x_gender
- ka_open

t-test for difference in means between alliance_US and alliance_RU

```{r}
fdi <- read.csv("fdi_data.csv")
dim(fdi)
head(fdi)
colnames(fdi)

### Key Variables


```

**2) Scraping More Info on S&P 500 Companies** - show appropriate use of at least ONE of each of the following – boxplot,
scatterplot (can be matrix plot), normal quantile plot (can be related to regression),
residual plots, histogram.

```{r}
# create scatterplot of gdp per capita vs. democracy index
plot(fdi$gdp_per_capita, fdi$democracy_index)
```

**3) Basic Tests** - t-test, correlation, AND ability to create bootstrap confidence interval for
either a t-test or a correlation.

```{r}
fdi_t_test <- fdi[,c("fdi_per_gdp", "alliance_US")]
t.test(fdi$fdi_per_gdp ~ fdi$alliance_US)


```

**4) Permutation Test** - include at least one.

**5) Multiple Regression** - use either backwards stepwise regression or some form of best
subsets regression. Should include residual plots. A GLM with a mix of continuous and
categorical predictors is fine here.

**6) AT LEAST ONE OF THE FOLLOWING TECHNIQUES** - ANOVA, ANCOVA, Logistic
Regression, Multinomial Regression, OR data scraping off a website.

```{r}
fdi_multinomial = fdi[]
```

#### Conclusion and Summary
